---
title: "Organizing Your Research Workflow with Rstudio"
subtitle: "A Practical Guide to Save You From Hating Yourself Later"
author: "Eli Kravitz"
date: "2/28/2019"
output:
  beamer_presentation:
    includes:
      in_header: preamble.txt
    keep_tex: yes
    slide_level: 3
    theme: metropolis
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("here")
```

### Philosphy

> "Always code as if the person who ends up maintaining your code is a violent psychopath who knows where you live." - CodeForTheMaintainer

* Usually this person will be you, so they will at least know where you live.



### Introduction

* I will give an \alert{opinionated} way for you to manage your research workflow in Rsudio.
  - I've learned this through trial and error... mostly error.

* Take parts of this you like, ignore parts you don't
  
* Many people are more qualified to talk about this
  - none of them will come here and do it for free. You're stuck with me



### Scenarios (totally not based on personal experience)

* __Scenarios 1__: You work on a project, complete 90% of it, then stop working on it for 2 years. 
  - \alert{Nothing works} on your new computer, you can't locate all the data and files you need. 
  
* __Scenarios 2__: You have a simulation working perfectly. You make a lot of changes. \alert{A few days later} you can't remember why you made these changes, and now \alert{your simulation does not work correctly}.

* __Scenarios 3__: You decide to rewrite most of the paper, show it to your adviser, and they tell you it was better before.
  - You only have the old draft of the paper in PDF. You DO NOT want to retype it.
  
### Storing Everything: Rprojects

* I organize all my research projects the same way.
   - Show folder structure
   
* Make the directory a Rproject.

* R projects let you:
   - Divide your work naturally
   - Store environments for every project
   - "freeze" R in a state that works (advanced, won't cover)
   
* Integrates nicely other things we'll cover

### Making an R project

* Go to File $\rightarrow$ New Project

\begin{center}
  \includegraphics[width=\textwidth,height= .90\textheight,keepaspectratio]{projects_new.png}
\end{center}

### Making an R project

* Follow the prompts to make your project.

* You're done. Easy.

* Now do this \alert{every time} you start a new research project.


### Relative Filepaths and here()

* Does your R code start with this?
```{r eval=F, echo=T}
setwd("C:\Users\Eli\path\that\only\I\have")
mydata = read.csv(data.csv)
```

* This is \alert{bad}. As soon as you rename or move directories, it breaks.

* What if someone else wants to use your code? What if you want to use a different computer? 

### Relative Filepaths and here()

* Putting your files in an Rproject already helps. The directory is to to where the .Rproj file is.

```{r echo=T}
getwd()
```

* Let's do better with the here package.

### Relative Filepaths and here()

* Here looks for a .Rproj file and makes file paths relative
```{r echo=T}
here()
here("Inner_Directory")
```
* Code won't need to change between computers

* Example: Load data
```{r echo = TRUE, eval=FALSE}
my_data = here("Data", "my_file.csv")
```


### Saving Progress with Github
* Now nothing is tied to the filepaths on our computer.

* How can we \alert{back everything up} and \alert{make it available on multiple computers}?

* Github!

### Saving Progress with Github

* Setting up Github is simple, but many steps.
  - See links at the end for instructions
  
* I will go over what it can do for you, and why you should work with it.

### Saving Progress with Github

* Github let's you track versions on your file

* Github saves the history of your file

* Made changes to your code $\rightarrow$ you can \alert{revert back to an earlier}. version.

* Works with .tex too!

* Show example on Github (code and .tex)

### Saving Progress with Github

* Github integrates really well with Rstudio
   - Comes pre-installed on current version of Rstudio
   
* Point-and-click integration

* Use this and make your life easier

### Make your presentations in Rstudio

* What else can we do in Rstudio? \alert{Presentations} in Rmarkdown

* Beamer is the worst. Does anyone find this intuitive? 
```{r echo = TRUE, eval=FALSE}
\begin{itemize}
\item
  I organize all my research projects the same way.

  \begin{itemize}
  \tightlist
  \item
    Show folder structure
  \end{itemize}
\item
  Make the directory a Rproject.
\item
  R projects let you:
```

### Make your presentations in Rstudio

* Look at how much more readable this is:

```{r echo = TRUE, eval=FALSE}
### Slide Title

* Bullet
  - Sub bullet
  
* Bullet

```

### Make your presentations in Rstudio

* Compare tables

* Latex:
```{r echo = TRUE, eval=FALSE}
\begin{table}
\begin{center}
  \begin{tabular}{ | l | c | r }
    \hline
    1 & 2 & 3 \\ \hline
    4 & 5 & 6 \\ \hline
    7 & 8 & 9 \\
    \hline
  \end{tabular}
\end{center}
\end{table}
```

### Make your presentations in Rstudio

* Rmarkdown:

```{r eval = F, echo=TRUE}
Table Header  | Second Header | Third Header
------------- | -------------- -------------
Table Cell    | Cell 2        |  Cell 3  
Cell 4        | Cell 45       |  Cell 6 
```

### Make your presentations in Rstudio

* You can create .tex files too. Just add to beginning of Rmarkdown

```{r eval = FALSE, echo = TRUE}
output:
  beamer_presentation:
    includes:
    keep_tex: yes
```

### DoRNG and Reproducability

* Many of you will need to do parallel computing for your research.

* Setting seeds is tricky

* Draw two examples on board

### DoRNG and Reproducability

* Fix this with DoParallel and DoRNG

* Automatically changes RNG type, feeds correct RNG to workers

* Saves the \alert{seeds as attribute}.
  - You can recreate these results!
  
 
### DoRNG and Reproducability
 
 * So easy, no reason not to do this
 
 * Setup:
```{r eval = F, echo = TRUE}
library("parallel"); library("doParallel"); library("doRNG")
cl = makeForkCluster(nnodes = detectCores()-1)
registerDoParallel(cl)
registerDoRNG(seed = 123)
```
 
### DoRNG and Reproducability

* Running reproducable parallel computing (easy too!)
```{r eval = F, echo = TRUE}
results = foreach(i = 1:Nsim) %dorng% {
  thing_I_want_in_parallel()
}
```

* Saves seed as attribute:

```{r echo = FALSE, message=FALSE, warning=FALSE, message=FALSE}
library("parallel"); library("doParallel"); library("doRNG")
cl = makeForkCluster(nnodes = detectCores()-1)
registerDoParallel(cl)
results = foreach(i = 1:2) %dorng% {
  rnorm(3)
}
```

```{r eval = TRUE, echo = TRUE}
attr(results, which = "rng")
```


### Resources

* I did not go over anything in detail, why anything works

* I wanted to demonstrate *why* you want to use these tools

* Next slides have resources for implementing all of this. They can teach you better than I can.



### Resources

* [Rprojects](https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects)

* [Github](https://guides.github.com/activities/hello-world/)

* [Github and Rstudio](https://jennybc.github.io/2014-05-12-ubc/ubc-r/session03_git.html)

* [Here function](https://github.com/jennybc/here_here)

### Resources

* [Presentations in Rmarkdown](http://svmiller.com/blog/2015/02/moving-from-beamer-to-r-markdown/)
   - I really, really like this website
   
* [Reproducable Parallel](https://cran.r-project.org/web/packages/doRNG/vignettes/doRNG.pdf)

